---
title: 'Exam 2'
author: "BENJAMIN OSEI-BLESSEN"
date: "DATA 2401 - Dr. Shoemaker"
output:
  html_document:
    fig_height: 3
    fig_width: 5
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = T, eval = T) # sets all code to show and run
```

# Instructions 

This exam is covered by the **UHD honor code**. Turning in this exam acknowledges that all work submitted on this exam is
your own, and that you have not used any off limit resource. You may use the R help files as needed. 



Submit your rmd *and* your knitted results, named `E2_LastNameFirstInitial.Rmd` and `E2_LastNameFirstInitial.html`, to Blackboard as instructed on the site. 

Code clarity and organization will be 5 points of your total grade. Make sure you have comments where needed and your code is easy to follow (and grade), do not print out whole dataframes. Please read this document carefully to make sure that you don't miss any requested task. 





#  Bills! 

I randomly generate some data of IDs and how much they owe. The second dataframe is a database of the ID numbers from payments that were made. 

```{r}
# this makes the randomness be the same way every time
set.seed(15) # leave it here

# makes a data frame of the ID and their amount owed
bills <- data.frame(id = 111:125, amount_owed = round(rnorm(15, 500, 20),2))
glimpse(bills)

# makes a data frame of id's the people who have made their payment 
  # (we're just sampling from the full list of id's, pretend it's the people who sent in the money)
who_paid <- data.frame(id = sample(111:125, size = 13))
glimpse(who_paid)
```


1. (5 points) You're in charge of finding the ID numbers of customers who haven't paid their bill. Use the appropriate join to do so. 

```{r}
anti_join(bills, who_paid)
```

+ 5 Extra Credit: Continue with your previous result; find the total amount that the fake company is still owed. 
```{r}
amt_owed <- anti_join(bills, who_paid)
sum(amt_owed$amount_owed)
```

#  Houses! 

Read in the houses.csv file from my GitHub using the below command. A subset of the information provided by the data owner is available in the appendix of this document.

```{r}
original_data <- read.csv("https://raw.githubusercontent.com/kshoemaker/Class_test/master/housedata.csv")
glimpse(original_data)
colnames(original_data)
```

2. (3 points) Create a new dataframe, with only the following variables: `SalePrice, YrSold, MoSold, ScreenPorch, Fireplaces, GrLivArea, CentralAir, YearBuilt, BldgType, LotArea, 'X1stFlrSF', PoolArea'` and `OverallCond`. 

```{r}
new_data <- original_data %>% 
  select(SalePrice, YrSold, MoSold, ScreenPorch, Fireplaces, GrLivArea, CentralAir, YearBuilt, BldgType, LotArea, X1stFlrSF, PoolArea, OverallCond)
new_data
```

Using an appropriate `dplyr` tool, rename `ScreenPorch` to `PorchSqFeet`, `GrLivArea` to `TotalSqFeet`, and `X1stFlrSF` to `SqFeetFirst` in your data frame.



```{r}
new_data <- new_data %>% 
  rename(PorchSqFeet = ScreenPorch, TotalSqFeet = GrLivArea, SqFeetFirst = X1stFlrSF)
glimpse(new_data)
```


3. (3 points) Is this data `tidy`? Why or why not? 

# It is Tidy. This is because every row represent an observation, each variable as it's own column, and each value as it's own cell.

4. (3 points) What types of variables does your new data frame have? 

# It has Categorical variables

## General Data Exploration

5. (15 points) Use `dplyr` to answer the below questions: 

- How old is the oldest house? 
- How much did the biggest house, by squarefeet, sell for? 
- What is the average total square footage and minimum lot area  for each building type? 
- How many of the houses have central air? 

```{r}
# How old is the oldest house? 
new_data %>%
  select(YearBuilt) %>% arrange(YearBuilt) %>% 
  summarise(oldest_house = 2021 - min(YearBuilt))

# How much did the biggest house, by squarefeet, sell for? 

#new_data %>% arrange(desc(LotArea)) %>% slice_max(LotArea) %>% select(SalePrice) 
new_data %>% select(SalePrice, LotArea) %>% 
  arrange(desc(LotArea)) %>% slice_max(LotArea)

# What is the average total square footage and minimum lot area  for each building type?
new_data %>% group_by(BldgType) %>% 
  summarise(avg_square_ft = mean(LotArea), min_lot_area = min(LotArea))

# How many of the houses have central air? 
new_data %>% select(CentralAir)  %>% filter(CentralAir == "Y") %>% count()

```



## Square feet and price

6. (5 points) Make a scatterplot of `TotalSqFeet` on the x-axis and `SalePrice` on the y-axis.

```{r}
new_data %>% ggplot(aes(x = TotalSqFeet, y= SalePrice )) + geom_point()
```

7. (5 points) Now, color the points by `BldgType`. This is a little too much to see clearly, no? 

```{r}
new_data %>% ggplot(aes(x = TotalSqFeet, y= SalePrice, color = `BldgType`)) + geom_point()
```

8. (5 points) Try again, but now facet by `BldgType`. 

```{r}
new_data %>% ggplot(aes(x = TotalSqFeet, y= SalePrice)) + geom_point() + facet_wrap("BldgType")
```

9. (5 points) What can you conclude from this plot? 

# From the plot, I cann tell that the price for a 1Fam houses where significantly higher than the remaining group.
# I can also tell that most of the observation are concentrated around 1000sqfts homes to 2000sqfts, except for "1fam" building type which are midly spread over to 3000sqfts and above.
# Finally I can read from the plot that for "1Fam" households, there is a positive relation between Salesprice and TotalSqft

## Porches

10. (5 points) Make a histogram of the porch square footage. Use `ggplot` instead of base plotting with `hist()`.

```{r}
new_data %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = PorchSqFeet))
```

11. (3 points) Create a new variable in your dataframe that indicates whether or not the house has a screened porch, something like T/F or "yes"/"no". How many houses have screened porches? How many don't? 

```{r}
#Creating new columns
new_data %>% select(PorchSqFeet) %>% mutate(porch_houses = PorchSqFeet>0)

#Number of houses with porch
new_data %>% select(PorchSqFeet) %>% filter(PorchSqFeet>0) %>% count()

#Number of houses without porch
new_data %>% select(PorchSqFeet) %>% filter(PorchSqFeet==0) %>% count()
```

12. (5 points) Filter the data for your histogram of the porch square footage to *only* houses *with* screened porches. 

```{r}
new_data %>% select(PorchSqFeet) %>% filter(PorchSqFeet>0) %>% 
  ggplot() + 
  geom_histogram(mapping = aes(x = PorchSqFeet))
```


13. (5 points) Color the points in the price vs. total squarefeet scatterplot by whether or not the house has a porch. You might consider making the points slightly transparent. 

```{r}

new_data %>% mutate(porch_houses = PorchSqFeet>0) %>% 
  ggplot(aes(x = TotalSqFeet, y= SalePrice, color = `porch_houses`)) + 
  geom_point(alpha=0.4)
```

14. (3 points) What, if anything, can you conclude from this? 

# After making it transparent, I realized that majority of the houses do no have a porch an so this plot does not really give me the information I need since most of the houses with porch are covered by the points of the houses without plot. 
 
15. (5 points) Are there any duplexes or two family conversions ("2fmCon") that have porches? If so, use the data and dplyr to display **only** their `SalePrice`, `TotalSqFeet`, and Porch squarefeet, and sort them from largest to smallest home square feet. 

```{r}
# No there is not
new_data %>% filter(BldgType %in% c("2FmCon", "Duplex")) %>%  
  select(PorchSqFeet, TotalSqFeet, SalePrice) %>% 
  mutate(porch_houses = PorchSqFeet>0) %>% 
  filter(porch_houses == "TRUE")

```


## Central Air is New?

16. (5 points) It would be reasonable to assume that newer houses have central air while older ones don't.  To explore this and see if the data supports this, plot an appropriate type of graph to look at YearBuilt, split by houses that do and don't have central air.  

```{r}
#new_data %>% select(YearBuilt)
  #ggplot(aes(y = YearBuilt, x = BldgType)) + 
  #geom_bar()
```

17. (5 points) There are several possible plot types for this question. What is another plot you could have used to explore the same question?  Plot it here. 

```{r}
# The condition and sale price
new_data %>% ggplot(aes(x = OverallCond, y= SalePrice)) + geom_col()

```


##  Your turn

18. (15 points) Using this home data, state a question and create a plot (or plots) of your own to answer that question. You may use any of the variables in our subsetted dataframe, or the 81 variables available in the full dataset. 

```{r}
# The condition and sale price
new_data %>% ggplot(aes(x = OverallCond, y= SalePrice)) + geom_col()

# 
```

Your question doesn't have to have an *interesting* answer, but your plot must answer the question. 

 

\newpage




## Appendix

This dataset is information about houses sold in a certain area from 2006 to 2010. The info for the variables I selected is below, more info can be found here: https://github.com/kshoemaker/Class_test/blob/master/data_description.txt
 
 
* `SalePrice` - the sale price of the house. 
* `YrSold` - year sold 
* `MoSold` - month sold
* `ScreenPorch` - Screen porch area in square feet
* `Fireplaces` - Number of fireplaces
* `GrLivArea` - above-ground living area square feet
* `CentralAir` - Central air conditioning
* `YearBuilt` - Original construction date
* `BldgType`
  +   `1Fam`:	Single-family Detached	
  +   `2FmCon`:	Two-family Conversion; originally built as one-family dwelling
  +   `Duplx`:	Duplex
  +   `TwnhsE`:	Townhouse End Unit
  +   `TwnhsI`:	Townhouse Inside Unit
* `LotArea` -  Lot size in square feet
* `PoolArea` - Pool area in square feet
* `OverallCon.` - Rates the overall condition of the house, 10 is "very excellent" to 1 is "very poor"