---
title: "Join Lab"
author: "Dr. Shoemaker"
date: "DATA 2401"
output: html_document
---

Case Study: Lahman’s Baseball Database

*Adapted from https://rpubs.com/williamsurles/293454. More advanced work is also available at https://rstudio-pubs-static.s3.amazonaws.com/390062_b835662a6f3b419eb7841165956eb3eb.html#lahman%E2%80%99s_baseball_database*

```{r}
library(Lahman) # you may need to install.packages() this
library(tidyverse)
```

Sean Lahman’s baseball database:
- 26 tables
- A great example of real world relational data

What are the universal variables? The keys?  Do any variable names span all of the datasets?





### Common keys?

There is a LOT of data here. We'll talk more about common keys when we talk about SQL. For now, you need to know that each player has a unique `playerID` that we'll be using to join. 



### Salaries

Who are the players? We can get list of distinct players, call the data frame `players`, by piping to `distinct`. 

How can we count how many players there are?

```{r}
players <- People %>% distinct(playerID)
nrow(players)

```


### Missing salaries

How many players are completely missing salary information? We can look at the dataframe `Salaries` to see. 

```{r}
anti_join(players, Salaries)
glimpse(Salaries)
```


We need to find the difference! Do an anti-join and count! (Remember our `by` variable is `playerID`)

Remember than an anti-join finds the members of dataframe A that are NOT in dataframe B. So, the anti-join of `players` with `Salaries` will give us the `playerID`s that are in the `players` database but NOT in the `Salaries` database. 

```{r}
anti_join(People, Salaries) %>% count()

```

### Unpaid games?

Is it possible that these players somehow did not play (and hence did not earn a salary)? We need to do a `semi-join` with `Appearances`, after our anti-join from above and count again! 

```{r}
head(Appearances)
```

```{r}
anti_join(players, Salaries) %>% semi_join(Appearances)
```

### How many games?

Can you determine how many games each of these unsalaried players played?

 - do our anti and left join from before
 - group by playerID
 - summarize by `sum`ming `G_all`
 - arrange in descending order 
 
```{r}
anti_join(players, Salaries) %>% 
        left_join(Appearances) %>% 
        group_by(playerID) %>% 
        summarise(sum_gall = sum(G_all)) %>% arrange(desc(sum_gall))
```


### How many at-bats?

Is it possible that the unsalaried players did not actually play in the games that they appeared in? Do our anti-join with salary, then a `left_join` with `Batting`, then sum their at bats, `AB`,  to find out.

```{r}
head(Batting)
```

```{r}
anti_join(players, Salaries) %>% 
        left_join(Batting) %>% 
        summarise(sum(AB))

```


### Hall of fame nominations

Let’s see how many players have been nominated for the Hall of Fame.

```{r}
head(HallOfFame)
anti_join(players, HallOfFame) %>% count()
```

Find (count) the distinct players that appear in HallOfFame. Create a new dataframe with the distinct players, then count that. 

```{r}
players_hof <- HallOfFame %>% distinct(playerID)
nrow(players_hof)
```

Join this distinct dataframe with the People data to get all the info about those players. 

```{r}
left_join(players_hof, People)
```

```{r}

```



There is SO MUCH more on the two websites listed above, if you're interested (or if you come back to these notes later when you're working on your project... )



