install.packages("datasauRus")
install.packages("tidyverse")

---
title: "R Project 1 - Hello R"
author: "Benjamin Osei-Blessen"
date: "Due sept 8th"
output: html_document
---

MAKE SURE YOU READ AND FOLLOW **ALL** INSTRUCTIONS ON THE INSTRUCTION DOCUMENT. 

### Load packages

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
```


### Exercise 1

1. From the Help file, how many rows and columns does the datasaurus_dozen file have? 1846. 142 rows and 13 columns.

2.  What are the variables included in the data frame? 3 variables: Dataset, x-values, y-values.


```{r}
datasaurus_dozen %>%
  count(dataset) %>%
  print(13)
```

### Exercise 2

First let's plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")
ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

And next calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

### Exercise 3

```{r plot-star}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()

```

for the correlation;
 
```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```

### Exercise 4

```{r}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()

```

```{r}


```

### Exercise 5

(Add code and narrative as needed. To add R chunks either type out the backticks, curly braces, and the letter `r` or use the Insert chunk button above, with the green +C.)
