---
output: 
  html_document: 
    fig_width: 5
    fig_height: 6
    highlight: tango
    theme: journal
---
install.packages("datasauRus")
install.packages("tidyverse")

---
title: "R Project 1 - Hello R"
author: "Benjamin Osei-Blessen"
date: "Due sept 8th"
output: html_document
---

MAKE SURE YOU READ AND FOLLOW **ALL** INSTRUCTIONS ON THE INSTRUCTION DOCUMENT. 

### Load packages

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
```


### Exercise 1

1. From the Help file, how many rows and columns does the datasaurus_dozen file have? 1846. 142 rows and 13 columns.

2.  What are the variables included in the data frame? 3 variables: Dataset, x-values, y-values.


```{r}
datasaurus_dozen %>%
  count(dataset) %>%
  print(13)
```

### Exercise 2

First let's plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")
ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

And next calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```

### Exercise 3

```{r plot-star}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()

```

for the correlation;
 
```{r cor-star}
star_data %>%
  summarize(r = cor(x, y))
```

### Exercise 4

```{r}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()

```

For the correlation;
```{r}
circle_data %>%
  summarize(r = cor(x, y))

```

### Exercise 5

```{r}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset))+
geom_point()+
facet_wrap(~ dataset, ncol = 3) + # facet the dataset variable, in 3 cols
theme(legend.position = "none")

```
```{r}
datasaurus_dozen %>%
group_by(dataset) %>%
summarize(r = cor(x, y)) %>%
print(13)
```

